---

title       : Computing Returns  
subtitle    :  
author      : Wilson Freitas  
job         : Quant  
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}  
highlighter : highlight.js  # {highlight.js, prettify, highlight}  
hitheme     : tomorrow      #   
widgets     : [mathjax]     # {mathjax, quiz, bootstrap}  
mode        : selfcontained # {standalone, draft, selfcontained}  

---

```{r setup, include=FALSE}
options(digits=4, width=80)
```

## download resources

- [PETR4.SA.csv](data/PETR4.SA.csv) file
- [index.Rmd](index.Rmd) source file

---

## loading time series into data.frame

```{r}
petr4.df = read.csv(file="data/PETR4.SA.csv", header=TRUE, stringsAsFactors=FALSE)
class(petr4.df)
str(petr4.df)
```

---

## data.frame `head`

```{r}
head(petr4.df)
```

---

## data.frame `tail`

```{r}
tail(petr4.df)
```

---

## data.frame properties

```{r}
colnames(petr4.df)
class(petr4.df$Date)
class(petr4.df$Adj.Close)
```

---

## indexing data.frame - always returns a vector

```{r}
petr4.df[1:5, "Adj.Close"] # using column name
petr4.df[1:5, 7] # using column index
petr4.df$Adj.Close[1:5] # using attribite reffering to column
```

---

## `drop=FALSE` keeps data.frame structure

```{r}
petr4.df[1:5, "Adj.Close", drop=FALSE]
```

---

## extracting data.frame subsets isn't easy

```{r}
c(which(petr4.df$Date == "2010-01-04"), which(petr4.df == "2010-07-01"))
petr4.df[which(petr4.df == "2010-01-04"):which(petr4.df == "2010-07-01"),]
```

---

## defining rownames

```{r}
petr4Prices.df = petr4.df[, "Adj.Close", drop=FALSE]
rownames(petr4Prices.df) = petr4.df$Date
head(petr4Prices.df)
```

---

## indexing by rownames

```{r}
petr4Prices.df["2010-01-04", 1]
petr4Prices.df["2010-01-04", 1, drop=FALSE]
```

---

## plotting data

```{r}
plot(petr4.df$Adj.Close)
```

---

```{r}
plot(petr4.df$Adj.Close, type="l", col="blue", lwd=2, ylab="Adjusted close", main="Monthly closing price of PETR4")
legend(x="topleft", legend="PETR4", lty=1, lwd=2, col="blue")
```

---

## simple returns

$$
R_t = \frac{P_t - P_{t-1}}{P_{t-1}}
$$

```{r}
n = nrow(petr4Prices.df) # number of rows
petr4.ret = (petr4Prices.df[2:n,1] - petr4Prices.df[1:(n-1),1]) / petr4Prices.df[1:(n-1),1]
head(petr4.ret)
```

---

### `petr4.ret` is not a data.frame object

```{r}
class(petr4.ret)
names(petr4.ret) = rownames(petr4Prices.df)[2:n] # adding names to a vector
head(petr4.ret)
```
---

### simple returns as data.frame objects - use `drop=FALSE`

```{r}
petr4.ret.df = (petr4Prices.df[2:n,1,drop=FALSE] - petr4Prices.df[1:(n-1),1,drop=FALSE]) / petr4Prices.df[1:(n-1),1,drop=FALSE]
head(petr4.ret.df)
```

---

## continuously compounded returns

$$
r_t = \log(1 + R_t) = \log \left( \frac{P_t}{P_{t-1}} \right)
$$

### easy way

```{r}
petr4.ccret = log(1 + petr4.ret)
head(petr4.ccret)
```
---

### alternatively

```{r}
petr4.ccret = log(petr4Prices.df[2:n,1]) - log(petr4Prices.df[1:(n-1),1])
names(petr4.ccret) = rownames(petr4Prices.df)[2:n]
head(petr4.ccret)
```

---

## comparing returns

```{r}
head(cbind(petr4.ret, petr4.ccret))
```

---

## plotting simple and continuous returns

```{r, eval=FALSE}
par(mfrow=c(2,1)) # split screen into 2 rows and 1 column
plot(petr4.ret, main="Monthly Simple Returns on PETR4", ylab="Return", type="l", col="blue", lwd=2)
abline(h=0)     
plot(petr4.ccret, main="Monthly Continuously Compounded Returns on PETR4", ylab="Return", type="l", col="blue", lwd=2)
abline(h=0)
par(mfrow=c(1,1)) # reset the screen to 1 row and 1 column
```

---

```{r, echo=FALSE}
par(mfrow=c(2,1)) # split screen into 2 rows and 1 column
plot(petr4.ret, type="l", col="blue", lwd=2, ylab="Return", 
	main="Monthly Simple Returns on PETR4")
abline(h=0)     
plot(petr4.ccret, type="l", col="blue", lwd=2, ylab="Return",
	main="Monthly Continuously Compounded Returns on PETR4")
abline(h=0)
par(mfrow=c(1,1)) # reset the screen to 1 row and 1 column
```

---

## on the same graph

```{r, eval=FALSE}
plot(petr4.ret, main="Monthly Returns on PETR4", ylab="Return", type="l", col="blue", lwd=2)
abline(h=0)     
lines(petr4.ccret, col="red", lwd=2)
legend(x="topright", legend=c("Simple", "CC"), 
	lty=1, lwd=2, col=c("blue","red"))
```

---

```{r, echo=FALSE}
plot(petr4.ret, main="Monthly Returns on PETR4", ylab="Return", type="l", col="blue", lwd=2)
abline(h=0)     
lines(petr4.ccret, col="red", lwd=2)
legend(x="topright", legend=c("Simple", "CC"), 
	lty=1, lwd=2, col=c("blue","red"))
```

---

```{r}
petr4.gret = 1 + petr4.ret # compute gross returns
petr4.fv = cumprod(petr4.gret) # compute future values
plot(petr4.fv, type="l", col="blue", lwd=2, ylab="$$", main="growth of $1 invested in PETR4")
```

---

## could life be easier?

R has a large number of packages that make life easier.

- [PerformanceAnalytics](http://cran.fhcrc.org/web/packages/PerformanceAnalytics/index.html): Econometric tools for performance and risk analysis
- [quantmod](http://cran.r-project.org/web/packages/quantmod/index.html): functions for quantitative modelling in finance
- [zoo](http://cran.fhcrc.org/web/packages/zoo/index.html): Z's ordered observations
- [xts](http://cran.fhcrc.org/web/packages/xts/index.html): Extensible time series
- [tseries](http://cran.fhcrc.org/web/packages/tseries/index.html): Time series analysis and computational finance
- [fPortfolio](http://cran.r-project.org/web/packages/fPortfolio/index.html): Portfolio Selection and Optimization
- [mvtnorm](http://cran.fhcrc.org/web/packages/mvtnorm/index.html): Multivariate normal and multivariate t distribution

---

## CRAN

<iframe src='http://cran.r-project.org/index.html' height="600px"></iframe>

---

## CRAN Task Views

<iframe src='http://cran.r-project.org/web/views/' height="600px"></iframe>

---

## Finance Task View

<iframe src='http://cran.r-project.org/web/views/Finance.html' height="600px"></iframe>

---

## PerformanceAnalytics package

```{r}
library(PerformanceAnalytics)
```

---

```{r}
dates.petr4 = as.yearmon(petr4.df$Date, format="%Y-%m-%d") # convert to month-year
head(dates.petr4)
petr4.z = zoo(x=petr4.df$Adj.Close, order.by=dates.petr4) # create zoo object
class(petr4.z)
head(petr4.z)
```

---

```{r}
plot(petr4.z, lwd=2, col="blue", ylab="Prices", xlab="Months")
```

---

## subsetting zoo objects

```{r}
petr4.z[as.yearmon(c("Jan 2010", "Jan 2011"))]
window(petr4.z, start=as.yearmon("Jan 2010"), end=as.yearmon("Jan 2011"))
```

---

## computing returns made simple

### core (no-package dependency)

```{r}
petr4Ret.z = diff(petr4.z)/lag(petr4.z, k=-1)
petr4Retcc.z = diff(log(petr4.z))
head(merge(petr4Ret.z, petr4Retcc.z))
```

---

### with PerformanceAnalytics functions

```{r}
petr4Ret.z = CalculateReturns(petr4.z, method="simple")
petr4Retcc.z = CalculateReturns(petr4.z, method="compound")
head(merge(petr4Ret.z, petr4Retcc.z))
```

```{r, echo=FALSE}
petr4Ret.z = diff(petr4.z)/lag(petr4.z, k=-1)
petr4Retcc.z = diff(log(petr4.z))
```

---

```{r}
chart.CumReturns(petr4Ret.z, lwd=2, col="red", main="Growth of $1")
```

---

## quantmod package

```{r}
library(quantmod)
```

---

```{r}
getSymbols("YHOO") # fetching time series as xts object
class(YHOO)
colnames(YHOO)
```

---

```{r}
start(YHOO) # first date
end(YHOO) # last date
```

---

```{r}
head(YHOO)
```

---

```{r}
chartSeries(YHOO, theme=chartTheme('white'))
```

--- &thanks
## Computing Returns