## Other approaches for creating time series objects

Instead of ordering `data.frame` object by its rownames we could simply use `xts` *constructor* to build time series objects from Yahoo! Finance CSV files.

First of all, load the CSV file using `read.csv` function.
This function returns a `data.frame` object and has some important parameters that might be used in order to make the work easier.

```{r, message=FALSE}
library(xts)
vale5.df <- read.csv('data/VALE5.monthly.raw.csv', header=TRUE, stringsAsFactors=FALSE, row.names=1)
```

The first *must-use* parameter is `header=TRUE` which tells that the very first line of the CSV file contains the columns names and that information is used to set `data.frame` columns names.
The second parameter is `stringsAsFactors=FALSE` which tells that string fields shouldn't be loaded as `factor` (another R primitive).
Factors are interesting when we have categorical data, that's not the case of date fields, so string dates must be loaded as strings.
The third parameter is `row.names=1` which sets the column 1 to `data.frame` row names.
It's quite important having the row names correctly defined because, once you have it defined properly you can simply call `xts` constructor passing the `data.frame` object and the row names are automatically set to time series index.
Once the time series is created it's ordered by its index attribute, so you have only to be sure of giving the data synchronized with its index.

```{r}
str(vale5.df)
head(vale5.df)
tail(vale5.df)
```

```{r}
vale5.ts <- as.xts(vale5.df)
str(vale5.ts)
head(vale5.ts)
tail(vale5.ts)
```

```{r}
as.yearmon(rownames(vale5.df), format="%Y-%m-%d")
vale5.ts <- as.xts(vale5.df, order.by=as.yearmon(rownames(vale5.df), format="%Y-%m-%d"))
str(vale5.ts)
head(vale5.ts)
tail(vale5.ts)
```
